### 2012-09-02 sections/implementation/flow-query-intermediate-format.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

A \ac{DNF} is a disjunction of conjunctive clauses. The elements of the
conjuctive clauses are terms. Each stage of the pipeline is represented in the
flowquery as a \ac{DNF} expression. The clauses in the \ac{DNF} are OR'd
together, while the terms in each clauses themselves are AND'd. The abstract
objects that store the JSON query and the results that incubate from each
stage are designed to be self-descriptive and hierarchically chainable. The
complete JSON query information for instance, is held within the
\texttt{flowquery} struct. Each individual branch of the flowquery itself is
described in a \texttt{branch} struct. A collection of these \texttt{branch}
structs are referenced in the parent \texttt{flowquery} struct.

>>>>

The entire JSON flow query is a collection of \ac{DNF} expressions. A \ac{DNF}
expression is a disjunction of conjunctive clauses. The elements of the
conjuctive clauses are terms. The clauses in the \ac{DNF} are OR'd together,
while the terms in each clause themselves are AND'd. The branchsets and each
\ac{DNF} expression of the pipeline stage is a \texttt{JSON array}.
\texttt{json-c} \cite{jsonc} is used to parse the flow query file. The
pipeline stages of the flow query language are also encapsulated in Python
classes and scripts have been written to facilitate the front-end parser to
serialize these pipeline objects to a \texttt{JSON} flow query.  The mapping
of the query to the structs defined in the execution engine is
not trivial. When reading the \texttt{JSON} query at runtime, the field
offsets of the NetFlow $v5$ record struct are read in as strings.  Utility
functions are defined that map the read field names to struct offsets and
field types and the operations to unique enum members.

The abstract objects that store the JSON query and the results that incubate
from each stage are designed to be self-descriptive and hierarchically
chainable.  The complete JSON query information for instance, is held within
the \texttt{flowquery} struct as shown in listing \ref{lst:struct}. Each
individual branch of the flowquery itself is described in a \texttt{branch}
struct.

\lstset{caption=Flow Query and Branch Structs,
				tabsize=2, language=C, numbers=left,stepnumber=1,
        basicstyle=\tiny\ttfamily, numberstyle=\ttfamily\color{gray},
        keywordstyle=\color{blue}, xleftmargin=15pt,
        rulesepcolor=\color{black}, label=lst:struct,
        aboveskip=10pt, captionpos=b}
\begin{lstlisting}
struct flowquery {
  size_t                                      num_branches;
  size_t                                      num_merger_clauses;

  struct branch**                             branchset;
  struct merger_clause**                      merger_clauseset;
  struct merger_result*                       merger_result;
  struct ungrouper_result*                    ungrouper_result;
};

struct branch {
  int                                         branch_id;
  struct ftio*                                ftio_out;
  struct ft_data*                             data;

  size_t                                      num_filter_clauses;
  size_t                                      num_grouper_clauses;
  size_t                                      num_aggr_clause_terms;
  size_t                                      num_groupfilter_clauses;

  struct filter_clause**                      filter_clauseset;
  struct grouper_clause**                     grouper_clauseset;
  struct aggr_term**                          aggr_clause_termset;
  struct groupfilter_clause**                 groupfilter_clauseset;

  struct filter_result*                       filter_result;
  struct grouper_result*                      grouper_result;
  struct groupfilter_result*                  gfilter_result;
};
\end{lstlisting}

The \texttt{JSON} query can also trigger and disable the stages at runtime.
This means that one only has to supply the constructs that one wishes to use.
The constructs that are not defined in the \texttt{JSON} query are inferred by
the execution engine as a disable request. The execution engine uses disable
flags that are turned on when the query is parsed. These flags are used
throughout the engine to only enable the requested functionality.

### 2012-09-02 sections/implementation/performance-optimizations.tex
* implementation   * nfql architecture (with figure)   * flow query intermediate format     * citation to json-c     * flow query and branch structs (listing)     * efficient rule processing using dedicated comparators (listing)   * splitter   * filter     * nfql vs nfql (optimized) (figure)   * grouper     * nfql (generic) vs nfql (specific) (figure)     * group aggregations merged in this sub section   * group filter, merger, ungrouper   * performance evaluations

There can be a situtation where the query designer may incorrectly ask for
aggregation on a field already specified in a grouper (or filter) clause. If
the relative operator is an equality comparison, the aggregation on such a
field becomes less useful, since the members of the grouped record will always
have the same value for that field. \ac{NFQL} realizes this redundant request
and ignores such aggregations.

@@

\ac{NFQL} pushes the filter stage out of the branch and closer to where the
trace is originally read. This enables the memory allocation of only records
that passed the filter stage thereby reducing the runtime memory footprint of
the execution engine.

Each branch runs in its own thread. If any of the stages of the branch return
a \texttt{NULL} when returning from their public interface function, there is
no reason to continue the thread. The subsequent stages of the branch cannot
do much with a \texttt{NULL} result.  Therefore, the branch thread returns
with a \texttt{EXIT\_FAILURE} if either stage returns \texttt{NULL}, and with
\texttt{EXIT\_SUCCESS} on normal exit.

Each stage of the processing pipeline is dependent on the result of the
previous one. As a result, the stages should only proceed and process, when
the previous returned results. Implementing such a response was
straightforward for the grouper and group filter, the merger although was a
little trickier.  The merger stage proceeds only when every branch has
non-zero filtered groups.  The iterator initializer
\texttt{iter\_init(\ldots)} deallocates and returns \texttt{NULL} if any one
branch has $0$ filtered groups.  Consequently a check is performed in the
merger to make sure \texttt{iter} is \emph{not} \texttt{NULL}.

>>>>

There can be a situation where the query designer may incorrectly ask for
aggregation on a field already specified in a grouper (or filter) clause. If
the relative operator is an equality comparison, the aggregation on such a
field becomes less useful, since the members of the grouped record will always
have the same value for that field. \ac{NFQL} realizes this redundant request
and ignores such aggregations.

@@

\ac{NFQL} pushes the filter stage out of the branch and closer to where the
trace is originally read. This reduces the memory allocation to only records
that passed the filter stage thereby reducing the memory footprint of the
execution engine.

Each stage of the processing pipeline is dependent on the result of the
previous one. As a result, the stages should only proceed, when the previous
returned results. Implementing such a response was straightforward for the
grouper and group filter, the merger although was a little trickier.  The
merger stage proceeds only when every branch has non-zero filtered groups.
The iterator initializer \texttt{iter\_init(\ldots)} deallocates and returns
\texttt{NULL} if any one branch has $0$ filtered groups.  Consequently a check
is performed in the merger to make sure \texttt{iter} is \emph{not}
\texttt{NULL}.
